---
title: "Introduction to `ggplot2` homework"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
    pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

# Homework Info

This homework recaps what you learnt in class today. 

You'll be looking at the `backpack` dataset contained within the `CodeClanData` package using tidyverse. You'll need to load the appropriate packages first.In all plots you make, ensure your labels are appropriate 

**You must write down a short summary of what each plot tells you in order to meet MVP.** Being an analyst is not just the ability to write code, you need to be able to interpret the visualisations and translate the relationships found (or not found) into plain language to explain to non-technical decision makers. Getting as much practice of this as you can while on this course is just (if not more!) important than the code!



# MVP

<br>

**Question 1.**  

Load in the libraries you need, and look at the `backpack` data. Take note of variable names and dimensions.
```{r}
library(tidyverse)
library(CodeClanData)
backpack
```



**Question 2.**

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.
```{r}
backpack %>% 
  ggplot(aes(backpack_weight)) +
  geom_histogram(aes(y = ..density..), colour = "lightgrey", fill = "orange", binwidth = 1) +
  geom_density(colour = "black")
```



**Question 3.**  

Are there differences in backpack weight between sex?
```{r}
backpack %>% 
  ggplot(aes(backpack_weight, colour = sex)) +
  geom_density()
```



**Question 4.**  

What is the relationship between body weight and backpack weight?
```{r}
backpack %>% 
  ggplot(aes(backpack_weight, body_weight)) +
  geom_point() +
  geom_smooth(se = F) +
  coord_flip()
```




**Question 5.**  

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.  
```{r}
backpack %>% 
  ggplot(aes(body_weight, backpack_weight, colour = sex)) +
  geom_point() +
  geom_smooth(se = F)
```

**Question 6.**  

What **proportion** of students report back pain?
```{r}
backpack %>%
  group_by(back_problems) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(x = "", y = count, fill = factor(back_problems))) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0(count, "%")), position = position_stack(vjust=0.5)) +
  theme_void() +
  labs(x = NULL, y = NULL, fill = "Back problems") +
  scale_fill_manual(values=c("#607A3C", "#E5FBC6"))
```



**Question 7.**  

Is this proportion different between men and women?
```{r}
backpack %>%
  group_by(back_problems, sex) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(
    x = sex, y = count, fill = factor(back_problems))) +
  geom_bar(stat = "identity") +
  geom_text(aes(
    label = paste0(count, "%")),
    position = position_stack(vjust=0.5)) +
  scale_fill_manual(values=c("#607A3C", "#E5FBC6")) +
  theme_minimal() +
  labs(x = "Sex", y = "Number", fill = "Back problems")
```



**Question 8.**  

Now you've investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both `backpack_weight` and `body_weight` as well as the `ratio`.
```{r}
backpack %>%
  mutate(wbp = sum(back_problems)) %>% 
  ggplot(aes(body_weight, backpack_weight, colour = factor(sex))) +
  geom_point() +
  geom_smooth(se = F) +
  facet_wrap(~back_problems)
```


*Hint: faceting would be useful here*.   

# Extensions

<br>

**Question 1.** 

Does reported backpain differ across the year groups? Ignore year 0 in your data.  


**Question 2.** 

Is there a difference between how heavy the bags those in different years carry?


