---
title: 'Manual model development'
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  pdf_document: default
---



# MVP

You are given a set of data on housing sale prices for the last few years in King County (near Seattle) between May 2014 and May 2015.

<br>
<div class="emphasis">
We want you to build an **explanatory model** for the `price` of housing in King County, i.e. an interpretable model in which the included variables are statistically justifiable.

The variable definitions are:

`id` - Unique ID for each home sold  
`date` - Date of the home sale  
`price` - Price of each home sold  
`bedrooms` - Number of bedrooms  
`bathrooms` - Number of bathrooms, where .5 accounts for a room with a toilet but no shower  
`sqft_living` - Square footage of the apartments interior living space  
`sqft_lot` - Square footage of the land space  
`floors` - Number of floors  
`waterfront` - A dummy variable for whether the apartment was overlooking the waterfront or not  
`view` - An index from 0 to 4 of how good the view of the property was  
`condition` - An index from 1 to 5 on the condition of the apartment  
`grade` - An index from 1 to 13, where 1-3 falls short of building construction and design, 7 has an average level of construction and design, and 11-13 have a high quality level of construction and design  
`sqft_above` - The square footage of the interior housing space that is above ground level  
`sqft_basement` - The square footage of the interior housing space that is below ground level  
`yr_built` - The year the house was initially built  
`yr_renovated` - The year of the house’s last renovation  
`zipcode` - What zipcode area the house is in  
`lat` - Latitude  
`long` - Longitude  
`sqft_living15` - The square footage of interior housing living space of the nearest 15 neighbors  
`sqft_lot15` - The square footage of the land lots of the nearest 15 neighbors  
</div>
<br>


# Question 1

Tidy up the data ready for regression:

```{r}
library(tidyverse)
houses <- read_csv("data/kc_house_data.csv") %>% 
  select(-id, -date, -sqft_living15, -sqft_lot15, -zipcode) %>% 
  mutate(waterfront = as.logical(waterfront),
         renovated = as.logical(yr_renovated),
         condition = factor(condition, levels = 1:5),
         grade = factor(case_when(grade > 10 ~ 1,
                                  grade > 7 ~ 2,
                                  grade == 7 ~ 3,
                                  grade > 3 ~ 4,
                                  TRUE ~ 5),
                        levels = 1:5)) %>% 
  select(-yr_renovated)
  

```

Following variable description I've converted grade into:

 - 11+ = high quality = 1
 - 8-10 = above average = 2
 - 7 = average = 3
 - 4-6 = below average = 4
 - 1-3 = low quality = 5


# Question 2

Check for aliased variables using the `alias()` function 

```{r}
alias(price ~ ., houses)
```


(this takes in a formula object and a data set). [**Hint** - formula `price ~ .` says 'price varying with all predictors', this is a suitable input to `alias()`].

Remove variables that lead to an alias. 

```{r}
houses <- houses %>% 
  select(-sqft_basement, -sqft_above)
```

```{r}
alias(price ~ ., houses)
```


# Question 3

Systematically build a regression model containing up to **four** main effects:

```{r, eval=FALSE}
library(GGally)
houses_tidy_numeric <- houses %>%
  select_if(is.numeric)

houses_tidy_nonnumeric <- houses %>%
  select_if(function(x) !is.numeric(x))

houses_tidy_nonnumeric$price <- houses$price

ggpairs(houses_tidy_numeric)
ggpairs(houses_tidy_nonnumeric)
```


and the same in subsequent rounds of predictor selection with the `resid` column.<br><br>
Remember, if you are not sure whether including a categorical predictor is statistically justified, run an `anova()` test passing in the models with- and without the categorical predictor and check the p-value of the test.
</details>

top numerical: sqft_living, view, lat:long (there will probably be an interaction), bedrooms, bathrooms
top categorical: waterfrontTRUE, grade1TRUE

There probably is a connection between view and waterfront.

```{r}
houses %>% 
  ggplot(aes(view, price, col = waterfront)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)
```

So view is another factor, and not numerical.
Doing some more tidying, eliminating some more variables

```{r}
houses <- houses %>%
  select(-waterfront) %>%
  mutate(view = factor(view, levels = 0:4))
```

```{r}
summary(lm(price ~ sqft_living + bedrooms, houses))
```

```{r}
anova(lm(price ~ sqft_living + bedrooms, houses), lm(price ~ sqft_living + bedrooms + bathrooms, houses))
```

After some testing it seems that sqft_living and bedrooms explain half of the variance.

```{r}
summary(lm(price ~ sqft_living + bedrooms + view, houses))
```



# Extensions

* Consider possible interactions between your four main effect predictors and test their effect upon $r^2$. Choose your best candidate interaction and visualise its effect.



* Calculate the relative importance of predictors from your best $4$-predictor model (i.e. the model without an interaction). Which predictor affects `price` most strongly?


sqft_living