---
title: "Homework - CIs"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```

<hr>

```{r include=FALSE}
library(tidyverse)
library(janitor)
library(infer)
```


# MVP

Now we'll go back to CI creation in the normal fashion. We'll take the `ames` data from the CIs lab earlier today and regard it now as a **sample**, we won't be drawing any smaller samples from within it. This is the usual situation in an analysis: you use all the data available to you!

<br>

**Task 1.**  
Load the data again, `clean_names()`, and re-familiarise yourself with it.
```{r include=FALSE}
ames <- read_csv("data/ames.csv") %>% clean_names()
```


<br>

**Task 2.**  
Investigate the distribution of `lot_area`. Is the distribution roughly normal? If not, what problems do you find?
```{r}
ames %>% 
  ggplot(aes(lot_area)) +
  geom_histogram(col = "white", bins = 50) +
  xlim(0, 100000)
  # The distribution is heavily right skewed
```



<br>

**Task 3.**  
Compute and visualise a bootstrap sampling distribution for the `mean(lot_area)` of the sold houses.
```{r}
obs_stat <- mean(ames$lot_area)

sampling_distribution <- ames %>% 
  specify(response = lot_area) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "mean")

sampling_distribution %>% 
  visualise() +
  shade_pvalue(obs_stat = obs_stat, direction = "none") # this is to visualise the observed sample mean, which is naturally around the centre of the simulated sampling distribution
```


<br>

**Task 4.**  
Use your bootstrap distribution to calculate a $95\%$ CI for `mean(lot_area)`, and visualise it on the distribution
```{r}
dist_ci <- sampling_distribution %>%
    get_ci(level = 0.95, type = "percentile")

sampling_distribution %>% 
  visualise() +
  shade_ci(endpoints = dist_ci)

```


<br>

**Task 5.**  
You would like to know the `mean(lot_area)` of the sold houses with higher confidence. Calculate the $99\%$ CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the $95\%$ CI? Does that make sense?
```{r}
ci_99 <- sampling_distribution %>%
    get_ci(level = 0.99, type = "percentile")

sampling_distribution %>% 
  visualise() +
  shade_ci(endpoints = ci_99)
```


<br>

**Task 6.**  
Calculate the point estimate of the `mean(lot_area)`
```{r}
obs_stat
```


<hr>

# Extension

<br>

**Task 1.**  
Calculate a point estimate and $95\%$ CI for the proportion of houses in the data built before 1920.  Does the number of `reps` you use matter? [Investigate `reps` from $200$ up to $50000$, memory of your laptop permitting].
<br><br>

```{r}
ames_1920 <- ames %>%
  mutate(before_1920 = as.numeric(year_built < 1920))

sampling_distribution <- ames_1920 %>% 
  specify(response = before_1920) %>% 
  generate(reps = 200, type = "bootstrap") %>% 
  calculate(stat = "mean")

ci_95 <- sampling_distribution %>%
    get_ci(level = 0.95, type = "percentile")

sampling_distribution %>% 
  visualise() +
  shade_ci(endpoints = ci_95)
```

